---
title: "Suplementary Code"
author: "Tess McCabe"
date: "April 6, 2015"
output: word_document
---

*This is the R code for the analysis of all ant data. It is broken into roughly three code blocks. The first section loads all of the nessisary packages and reads in the data. The second section is code for the figures displaying the descriptive statistics. The third section outlines the statistical methods used and include brief notes and visualizations of thier interpretation. All datasets used in these analyses are availible at Harvard Forest Datasets*




**Packages & Data**

```{r}
#### Packages
library(permute)
library(lattice)
library(vegan)
library(vegetarian)
library(car)
#Note that a package not published on the CRAN server "biostats.r" is sourced later. For more information please contact Kevin McGarigal of the University of Massachusetts, Amherst
#Initialization
#### Read in Data 
setwd("/Users/tessmccabe/Dropbox/Junior/Junior Year research/Ant Stats/data") #Set to appropriate local working directory
source('biostats.R')
antdatanum<-read.csv("ants.csv") #Only the ant portion of Harvard Forest dataset. All three "Q" Charachtors were removed manually. 
#Data Preparation
antdatanum$Aphpic<-antdatanum$Aphpic+antdatanum$Aphrud #Combine both Aphpic and Aphrud 
antdatanum<-antdatanum[,-6] # nix aphrud
antdata2 <- antdatanum[,-c(1:3)]#Create a matrix without lable columns
antdata2[antdata2>0] <-1 #Make that matrix presence vs absence change all numbers into 1's. 
antdata3<-cbind(antdatanum[,1:3],antdata2) #Create a presence vs absence matrix with lable columns. 

```


**Figures & Discriptive Statistics**
#Plotting
```{r, echo=FALSE}
#### Species Occurences 
antdata2Plot<-(c(sum(antdata2$Aphful),sum(antdata2$Aphpic),sum(antdata2$Camnov),sum(antdata2$Campen),sum(antdata2$Forase),sum(antdata2$Forneo),sum(antdata2$Forrub),sum(antdata2$Forsubi),sum(antdata2$Forsub),sum(antdata2$Lasali),sum(antdata2$Lasumb),sum(antdata2$Mryame),sum(antdata2$Myrpun),sum(antdata2$Steimp))) #a vector of the sum of every column
par(mar=(c(10, 4, 4, 2) + 0.1))# Set margins for barplot
barplot(antdata2Plot, ylab= "Occurrence", names.arg= c("Aphaenogaster fulva","Aphaenogaster picea/ rudis","Camponotus noveboracensis","Camponotus pennsylvanicus","Formica aserva","Formica neogagates","Formica rubicunda","Formica subintegra","Formica subsericea","Lasius alienus","Lasius umbratus","Myrmica americana","Myrmica punctiventris","Stenamma impar"), las=3, cex.names=0.8, main="Species Occurence")
par(mar=(c(5, 4, 4, 2) + 0.1)) #Reset paramaters for other figures 
#Plot of Species Occurences

#### Species Richness by Year
specRichYear1<-rowSums(antdata3[antdata3$Year=="2006",-c(1:3)])
specRichYear2<-rowSums(antdata3[antdata3$Year=="2007",-c(1:3)])
specRichYear3<-rowSums(antdata3[antdata3$Year=="2008",-c(1:3)])
specRichYear4<-rowSums(antdata3[antdata3$Year=="2009",-c(1:3)])

sdyear<-c(sd(specRichYear1),sd(specRichYear2),sd(specRichYear3),sd(specRichYear4))#Vector of standard deveations
barplot(c(mean(specRichYear1),mean(specRichYear2),mean(specRichYear3),mean(specRichYear4)), names.arg= c("2006", "2007", "2008","2009"), ylab="Average Species Richness", main= "Species Richness by Year") #Displays the mean number of species per year


#### Species Richness by Treatment
specRichTreat1<-rowSums(antdata3[antdata3$Treatment==1,-c(1:3)])
specRichTreat2<-rowSums(antdata3[antdata3$Treatment==2,-c(1:3)])
specRichTreat3<-rowSums(antdata3[antdata3$Treatment==3,-c(1:3)])
specRichTreat4<-rowSums(antdata3[antdata3$Treatment==4,-c(1:3)])

sdtreat<-c(sd(specRichTreat1),sd(specRichTreat2),sd(specRichTreat3),sd(specRichTreat4))# Vector of standard deveations
barplot(c(mean(specRichTreat1),mean(specRichTreat2),mean(specRichTreat3),mean(specRichTreat4)), names.arg= c("Control", "Warming", "Nitrogen","Nitrogen/Warming"), ylab="Average Species Richness", main= "Species Richness by Treatment", col=c("purple","orange","blue", "green"))#Displays the mean number of species per treatment 
```


**Statistical Tests**
#Data Analysis
Part one: Multivariate Analysis 
```{r}
#### Manova
rowZero<-antdata3[(rowSums(antdata3[,-c(1:3)])!=0),]# Check for and remove any rows without any species found. 
mv.outliers(rowZero[,-c(1:3)], method='bray')# Returns no outliers
adonis(rowZero[,-c(1:3)]~(rowZero$Treatment/rowZero$Plot)*rowZero$Year, method='bray')# a Non-Patametric MANOVA function.  Only found significant differences between different years. 


# Check to see if 2007 is the cuase of significant difference, by deleteing 2007. 
rowZero1<-rowZero[rowZero$Year!="2007",]# Removes 2007
adonis(rowZero1[,-c(1:3)]~rowZero1$Treatment/rowZero1$Plot*rowZero1$Year,method='bray')# Year is still significant. Results warrent more investigation. 
#Plotting
### Visual Confirmation of Manova resuls 
treat<-rowZero$Treatment
year<-rowZero$Year
yearBeta<-betadisper(vegdist(rowZero[,-c(1:3)], method= 'bray'), group= year, type=c("median","centroid")) #Creates a betadisper object useing a vector of bray distance measures  grouped by year
treatBeta<-betadisper(vegdist(rowZero[,-c(1:3)], method= 'bray'), group= treat, type=c("median","centroid")) #Creates a similar betadisper object grouped by treatment
plot(yearBeta, axes = c(1,2), cex = 0.7, hull = TRUE, main= "Year Visualization of NPMANOVA")
plot(treatBeta, axes = c(1,2), cex = 0.7, hull = TRUE, main= "Treatment Visualization of NPMANOVA") #The difference between the two plots is clear.The year group look more distinct.  Results of MANOVA likely accurate. 
#Data Analysis
### Pairwise Comparisons of years
permutest(yearBeta, pairwise =TRUE,control = how(nperm = 999)) #Only significant difference between 2007 & 2008 and 2008 & 2009. None are significantly different from control year.  

permutest(treatBeta, pairwise =TRUE,control = how(nperm = 999)) #A significant value between treatment 1 (control) and warming/nitrogen, as well as a just verging on significance value for control and just nitrogen. Because the adonis found no significant difference will disregard. 
```

Univariate Analysis
#Data Preparation
```{r}
rowZeroNum<-antdata3[(rowSums(antdata3[,-c(1:3)])!=0),] 
rich<-rowSums(rowZeroNum[,-c(1:3)]) #Creates a vector of the species richness for each plot
rich_aov<-aov(rich~rowZeroNum$Treatment*as.factor(rowZeroNum$Year)) #create an ANOVA object. 
#Plotting
plot(resid(rich_aov)) # Plot siduals to double check normality. Appears normal. 
Anova(rich_aov) #Year is the only significant term. 
```

Check of control year 2006 
```{r}
#Prep
antdata2006<-antdata3[which(antdata3$Year=="2006"),] # Subset of 2006
rowZero2006<-antdata2006[(rowSums(antdata2006[,-c(1:3)])!=0),] # remove any rows with zero values
#Plotting
rich2006<-rowSums(rowZero2006[,-c(1:3)]) #creates a richness value for all plots in 2006

aov2006<-aov(rich2006~as.factor(rowZero2006$Treatment)) #Makes an anova object
Anova(aov2006) #No significance found. 
```

Adjusted p-value of all stitistical tests
```{r}
p.vals<-c(0.49325,0.05188,0.67472, 0.298,0.007,0.991,0.672,0.810,0.09035,0.1197,.000005,.184,.2881,.000005)#A vector of all p-values to be adjusted. Includes values from rich_aov (3 values), and the adonis (4 values), and permutetest.yearbeta (6 values).  
p.adjust(p.vals, method='holm')
```
*Special thanks to group members Ashni, Carly & Lauren!*