---
title: 'Questions about GLM''s '
author: "Tess"
date: "February 16, 2015"
output: word_document
---

Hello, here's a summery of my struggle with GLM's. 

This section of code just reads in/ cleans up all of the data sets, the glm stuff starts in the second chunk of code. 

```{r, echo=TRUE}
library(permute)
library(lattice)
library(vegan)
library(vegetarian)
library(pscl)

setwd("/Users/tessmccabe/Dropbox/Junior/Junior Year research/Ant Stats/data")
antdatanum<-read.csv("ants.csv") #I set data as working directory
antdata2 <- antdatanum[,-c(1:3)]
# take all abundance data and replace it with 0s and 1s
antdata2[antdata2>0] <-1
# append columns with plot information to ant occurrence data
antdata3<-cbind(antdatanum[,1:3],antdata2)


```

Start of GLM experimmentation
Failed code: 
```{r, echo=TRUE}
##this code doesn't run. It's been commented out, but it should produce the same error messeges when it's run without comments. 
#glm(antdata3~Treatment*Year,family=poisson()) #Gives "invalid type (list)"" error
#glm(antdata2~Treatment*Year,family=poisson()) #same error
#glm(antdata3~antdata3$Treatment*antdata3$Year,family=poisson()) #again
#glm(antdata3$Aphful~antdata3$Treatment*antdata3$Year*antdata3$Aphpic*antdata3$Aphrud*antdata3$Camnov*antdata3$Campen*antdata3$Forase* antdata3$Forneo*antdata3$Forrub*antdata3$Forsubi*antdata3$Forsub*antdata3$Lasali*antdata3$Lasumb*antdata3$Mryame*antdata3$Myrpun*antdata3$Steimp,family=poisson()) #This does work, but gives the warning messege "glm.fit: fitted rates numerically 0 occurred" and only incorperates one species. 
#glm(antdata3$Aphful:antdata3$Steimp~antdata3$Treatment*antdata3$Year,family=poisson()) # ###Gives error "Error in model.frame.default(formula = antdata3$Aphful:antdata3$Steimp ~  : 
#  variable lengths differ (found for 'antdata3$Treatment')
#In addition: Warning messages:
#1: In antdata3$Aphful:antdata3$Steimp :
#  numerical expression has 96 elements: only the first used
#2: In antdata3$Aphful:antdata3$Steimp :
 # numerical expression has 96 elements: only the first used"
#DEADLY CODE: CAUSES  CRASHES
#did.it.work<-glm(antdata3$Aphful~antdata3$Treatment*antdata3$Year*antdata3$Aphpic*antdata3$Aphrud*antdata3$Camnov*antdata3$Campen*antdata3$Forase* antdata3$Forneo*antdata3$Forrub*antdata3$Forsubi*antdata3$Forsub*antdata3$Lasali*antdata3$Lasumb*antdata3$Mryame*antdata3$Myrpun*antdata3$Steimp,family=poisson()) 


```


Here's what worked: 

 #start by making a dataframe of form: plot# treatment# year# Species name Count#
Species<-c(antdata3$Aphful,antdata3$Aphpic,antdata3$Aphrud,antdata3$Camnov,antdata3$Campen,antdata3$Forase, antdata3$Forneo,antdata3$Forrub,antdata3$Forsubi,antdata3$Forsub,antdata3$Lasali,antdata3$Lasumb,antdata3$Mryame,antdata3$Myrpun,antdata3$Steimp)
#checked and this does have a lengthof 1440, or 15 * 96, or the number of elements in the matrix

Year<-antdata3[,1]
Treatment<-antdata3[,3]
glm.dataframe<-cbind(Year,Treatment,Species) #so that I can access the variable names
as.data.frame(glm.dataframe) # I have to coerce it. But it doesn't work if I just coerce it
#so I chop up the columns again
Year<-glm.dataframe[,1]
as.character(Year)
Species<-glm.dataframe[,3]
Treatment<-glm.dataframe[,2]
as.character(Treatment)
#now I can run the glm function
glm.test<-glm(Species~Treatment*Year, family=quasipoisson)
summary(glm.test)
plot(glm.test)

#still getting an AIC of nearly 900


This still lumps all of the species counts into one vector, which doesn't really allow for effect comparisons between species. Do you have any advice for how to format the data so that I can use it in the GLM function? 

Thanks, 


-Tess McCabe


Ok: here's the deal, now I'm fitting a glm for diversity indices alpha and beta

```{r, echo=TRUE}
#Let's start with making a vector of alpha  values
d(antdata2[c(3),], lev = "alpha", wts = FALSE, q = 1, boot = FALSE) # USed this to test that the alpha values vector reflects this function's values. Checked out. 
Alpha_values<-apply(antdata2,1,d)
#Alpha_values<-log(Alpha_values)
alpha_glm<-glm(Alpha_values~antdata3$Treatment*antdata3$Plot*antdata3$Year)
plot(alpha_glm)

summary(alpha_glm)
alpha_glm2<-glm(Alpha_values~antdata3$Treatment*antdata3$Year)
plot(alpha_glm2)
summary(alpha_glm2)


#Trying a zero inflated model 
#alpha_zeroinfl<- zeroinfl(Alpha_values ~ antdata3$Treatment*antdata3$Plot*antdata3$Year) # Not quite. 
```

Look up non-parametric tests for ancova
Play with data transformations

#methods from Frey paper: Harvard forest paper methods to prove that 

Eviroemtal data: 
```{r, echo=TRUE}
#rarefaction
#you want to report the error in observation. 

```