---
title: "Lab 8"
author: "Tess McCabe"
date: "April 8, 2015"
output: html_document
---

**Question 1**

```{r}
library(lattice)
library(permute)
library(car)
setwd("/Users/tessmccabe/Dropbox/Junior/Junior Year research/Ant Stats/data")

#### Begining of Lab 8
source('biostats.R') # Load functions in the informal biostats library

fishdata<-read.csv("Fish_data.csv")
#### In question 1 
table(fishdata$Treatment) #Determine the number of sample rows per Treatment
#All treatments have at least 1 

fish2<-subset(fishdata, fishdata$Year>2006)
#Make a subset of data

fish3<-fish2[,-c(1:4)]
fish3[fish3>0]<-1
##Created the incedence matrix 
###Part 3 Rowsums! 

Rich<-rowSums(fish3) #Has 30 variables, represents richness of each replication
fishbin<-Rich # Oops, I guess we're calling this somethign else
fish4<-cbind(fish3,fishbin)#Check, and it looks alright. 

```

**Question 2: Entering Anova land**

```{r}
richness_mod<-aov(fishbin~fish2$Treatment*fish2$Year)# Anova that looks at group differences of year, treatment, and thier interaction
plot(resid(richness_mod)) # The residuals actually look pretty good, if there's a bit of a bump in the middle/ or a band. 
#Downloaded car package

Anova(richness_mod) # Both Year and treatment are significant! 
#Pairwise compareisons
TukeyHSD(richness_mod) # WOW. SO SIGNIFICANT. But it didn't run for years, cuz  those aren't factors. That's fine though, because we just found out that year was significant. 

```
This was a very exciting analysis. Not only were both year and treatment significant, but each pairwise comparison of the treatment groups were also significantly different. The pruned-intact difference was less significantly different than either treatment with the cleared. 


**Ordination Sensation**

```{r}
#### Lab Manual Code

# Subset data into abundances per linear meter for intact, cleared, vs. pruned comparison years 2008 & 2015
cpi <- subset(fishdata, Year>2006)
cpi <- rbind(subset(cpi, Treatment=="pruned"),subset(cpi, Treatment=="intact"), subset(cpi, Treatment=="cleared"))

# Perform row summary statistics to see if there are any rows containing all zeros
sum.stats(cpi, var='CHCA:UNKN', margin='row')

# Remove rows with all zeros
cpi_no0rows <- cpi[rowSums(cpi[,5:dim(cpi)[2]])!=0, ]

# Perform column summary statistics to see if there are any columns containing all zeros
sum.stats(cpi_no0rows, var='CHCA:UNKN', margin='column')

# Remove columns with all missing values
cpi_no0 <- subset(cpi_no0rows, select=-c(CHST, PAIN, BLEN, THBI, CHMU, CTSA, GRLO, SPGU, TYCR, GOGR, HAPA, EUGU, SILV, TEUT))

# Create a response variable y that consists of the species abundance per linear meter data matrix
y <- cpi_no0[,5:dim(cpi_no0)[2]]

# Standardize with a column total standardization to put relative species abundance to a species profile
y.std <- data.stand(cpi_no0[,5:dim(cpi_no0)[2]], method='total') # why this step?
```



**Question 4**

```{r}
## Are there mv outliers? 
mv.outliers(y.std, method='bray') #This didn't return any outliers! -good sign. 

## is there colinearity 
splom(y.std) # This produces an image that isn't really easy to interpret. 
splom(y.std[1:10])# This is better, and shows non-linearity. A PCA isn't apropriate because it relies on linearity. 


## NMDS
#Lab manual code. 
# Create a scree plot to determine dimensionality (k) of the NMDS
nmds.scree(y.std, distance='bray', k=6, trymax=50, autotransform=FALSE)
#Based on this, 5 axis gets us at 0.05 stress. 
```
The first time stress dips below 0.05 is when there are five axes. We could go to six axes, but that wouldn't follow the principle of pasimony. 



**Question 5**

```{r}
## Lab manual code
# Conduct NMDS. Be sure to fill in a value for k! #Done :)
cpi_nmds <- metaMDS(y.std, distance='bray', k=5, trymax=50, autotransform=FALSE)

# Create stress plot of relationship between the original dissimilarities and Euclidean distances in the ordination species
stressplot(cpi_nmds)

```

Based on the plot, the fit is really good. The linear R2 value if 0.92 which is incredible. This mans that the equation generated really well represent the point in the plot. 


**Question 7**

```{r}
# Create a vector corresponding to the grouping variable (i.e., Treatment)
grp <- cpi_no0$Treatment #Making a smaller argument for the ordihull f(x)n

p <- ordiplot(cpi_nmds, choices=c(1,2), display='sites') #Makes a plot of all of the "cleaned up" matrix data

ordihull(p, groups=grp, col='blue', show.groups="cleared") # Draws the cleared group on the plot

ordihull(p, groups=grp, col='red', show.groups="intact") #Draws the intact group on the plot

ordihull(p, groups=grp, col='orange', show.groups="pruned") # Draws the pruned group on the plot. 

```
 This plot represents the three trials plotted in species space. 