---
title: "McCabe_Lab9"
author: "Tess McCabe"
date: "April 15, 2015"
output: html_document
---

Setup

```{r}
require(raster) #Version 2.3-24
require(sp) #Version 1.0-17
library(dismo) #Version 
library(gbm)
setwd("/Users/tessmccabe/Desktop")
source('biostats.r')
```



```{r, echo=FALSE}
##Lab Manual code
data(Anguilla_train) # Load the Anguilla australis data.
head(Anguilla_train) # Check out the column names and first several lines of the training data
```

Question 1 
```{r}
tail(Anguilla_train) # Final site number is 1000, I assume there are 1000 sites. This matches the 1000 obs in the summary. 

Atrain<-Anguilla_train[Anguilla_train$Angaus==1,] 
dim(Atrain) # Only 202 obs long. Only 202 of sites have presence recorded. 
```

Question 2: Fitting the BRT

```{r}
### Manual Code
anguas.tc5.lr005 <- gbm.step(data=Anguilla_train, gbm.x=3:13, gbm.y=2, family="bernoulli", tree.complexity=5, learning.rate=0.005, bag.fraction=0.5)
```

According to the graph, it takes about 1300 trees, and at that point the deviance is around 0.7. With more trees,  the devience value raises slightly. 


Question 3

The four values that the paper disscussed running were the learning rate, lr, the tree complexity, tc, the number of trees, nt, and the bag fraction. To determine which value of lr and nt was most apropriate, the authors plotted the devience against nt for different values of lr (from 0.005-0.05 at interavls). By looking at the different curves the plot made, the authors were able to tell that models that increades dramatically indevience after hitting a devience minimum were over fit, and they could choose the largest lr value that didn't seem to be overfit.  They compared these graphs for two different tc values, and then chose the one that offered the most aporpriate options of lr.

The Bag fraction is the other value that was disscussed. The defualt is 0.5 or 50% of the data used for training and 50% used for testing the model. They recommend running one's own sensistivity analysis by comparing predictive performence by fraction. They seem to have fount that a bag froaction of 0.5-0.75 is apropiate for presence vs absence data. 

Question 4

```{r}
summary(anguas.tc5.lr005) #Note that the y axis labels are terrible 

```
The six top predictive values were: 

SegSumT      
USNative    
Method        
DSDist       
DSMaxSlope 
USSlope       


Question 5

```{r}
gbm.plot(anguas.tc5.lr005, n.plot=6)
par(mfrow=c(1,1)) #Thanks Carly for fixing this
```
The segsum one shows that you are morelikely to find the eels during a warm summer thant diring a cold one.  USNative shows that areas with lower us native trees you are more likely to find an eel. The third graph shows that electric fishing is slightly more effective than trap/net/manual fishing.  The DSdist graph  shows that notmally the closer to the coast, the more likely you are to see eels (although there is a bump at 100- maybe anouthetr source of water?).  The DSMaxSlope means the maximum downstream slope, (Which the eels seem to hate), and finally the USSlope is the average upstream slope. Based on these, it looks like the eels are found in warm summers, around invasisve trees, close to the shore, in flat streams. 

Question 6

```{r}
### Maual code
par(mfrow=c(1,1)) #Thanks Carly for fixing this
data(Anguilla_grids) # Read in the RasterBrick of spatial data for the predictor variables
plot(Anguilla_grids) # Plot the rasters for the predictor variables

Method <- factor('electric', levels = levels(Anguilla_train$Method)) # Create a constant value of class factor for fishing method to pass on to the predict function
add <- data.frame(Method) # Turn the Method constant values into a data frame as that is the form it needs to be in for the *predict()* function from the raster package
p <- predict(Anguilla_grids, anguas.tc5.lr005, const=add, n.trees=anguas.tc5.lr005$gbm.call$best.trees, type="response") # Predict the BRT model using the spatial gridded raster data of predictor variables
p <- mask(p, raster(Anguilla_grids, 1)) # The mask function ensures that the plotting function knows not to plot NA values in the predicted response raster (i.e., plots white space between the rivers where we do not have spatial data)
plot(p, main='Angaus - BRT prediction') # Plot the predicted map based on the BRT model
```

It does look like it makes intutive sence.  The Segsum plot has a similar shape to the predictive model. Which does suggest that it is highly influenced by the Seg Sum variable. It seems a bit odd that the eels are more likely to be present in more areas than absent in more areas. 