---
title: "Lab2"
author: "Tess"
date: "February 4, 2015"
output: html_document
---


```{r}
setwd("/Users/tessmccabe/Dropbox/Junior/Spring/Computational Methods/Data")
spore_disp<-read.table("spore_disp.txt")
```

**Question 1: a)/b)**

The data is continuous, because both mass and distance are a continuous measures. For example, you can have 3 inches or 2.5 inches or 2.25 inches etc.  
Console

read.table("spore_disp.txt")
mass distance
1    2       20
2   10       18
3   15       14
4   45       10
5  160        1


**Question 2:**
```{r, echo=TRUE
plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("black",4),"red"))
```

Console

plot(spore_disp)
> plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0)
> plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("black",4),"red")
+ col=c(rep("black",4),"red")
Error: unexpected symbol in:
"plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("black",4),"red")
col"
> ```
Error: attempt to use zero-length variable name
> col=c(rep("black",4),"red")
> plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("black",4),"red"))


**Question 3: a)/b)**

rep() repeats the value of "black" 4 times as a shorthand for writing `r c("black","black","black","black","black","red")`. To make all of the points under 150 green I could just insert `r plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("green",4),"red"))`
```{r, echo=TRUE}
plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("green",4),"red"))
```

Console

 plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("green",4),"red"), yaxp=(0,20,20))
Error: unexpected ',' in "plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("green",4),"red"), yaxp=(0,"
> ?yaxp
No documentation for ‘yaxp’ in specified packages and libraries:
you could try ‘??yaxp’
> plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("green",4),"red"), yaxp=c(0,20,20))


**Question 4:**
```{r, echo=TRUE}
plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("green",4),"red"), yaxp=c(0,20,10))

```

Console

 legend(locator(1),legend=c("Mass<150 mg","Mass> 150 mg"), col=c("black", "red"))
Error in legend(locator(1), legend = c("Mass<150 mg", "Mass> 150 mg"),  : 
  invalid coordinate lengths
legend(locator(1),legend=c("Mass<150 mg","Mass> 150 mg"), col=c("black", "red"))
 ?legend
?locator
?legend


**Question 5**
```{r, echo=TRUE}
plot(spore_disp,main="Spore dispersal", xlab="Mass (mg)", ylab="Distance (m)",pch=0, col=c(rep("black",4),"red"), yaxp=c(0,20,10))
# The legend() puts a legend on the graph at whatever x/y you spesify.
# locator() is super neat because to saves the x/y cordinates of your mouse when you click. 
# legend= referes to a vector of all of the lines of text within the legend
# col refers to the color of the labes (also need a vector)
#legend(locator(1),legend=c("Mass<150 mg","Mass> 150 mg"), col=c("black", "red"))
#NOTE the legen code is commented out because RMarkdown doesn't like the locator function. 

```

```{r, echo=TRUE}
#legend modified to include symbols
#legend(locator(1),legend=c("Mass<150 mg","Mass> 150 mg"), col=c("black", "red"), pch=0)
#NOTE the legen code is commented out because RMarkdown doesn't like the locator function. 
```

Console

> legend(locator(1),legend=c("Mass<150 mg","Mass> 150 mg"), col=c("black", "red"), pch=0)
> legend(locator(1),legend=c("Mass<150 mg","Mass> 150 mg"), col=c("black", "red"))
Error in legend(locator(1), legend = c("Mass<150 mg", "Mass> 150 mg"),  : 
  invalid coordinate lengths



**Question 6:**
```{r,echo=TRUE}
#sample() referes to a funtion that selests onbjects from a vector
#seq() refers to a function that generates a sequence from the first number to the second number, itereating by the third number. 
#tempurature equals two vectors of 50 numbers, one with values between 41 and 5 and the other between the values of 105 and 75. 
tempurature<-c(sample(seq(5,41,1),50,replace=TRUE),sample(seq(75,105,1),50,replace=TRUE))
# season is a vector that has labled the virst half of the data "Winter" and the second half "summer"
season <- c(rep("winter",50), rep("summer",50))
```

Console

tempurature<-c(sample(seq(5,41,1),50,replace=TRUE),sample(seq(75,105,1),50,replace=TRUE))
season <- c(rep("winter",50), rep("summer",50))

**Question 7:**
```{r,echo=TRUE}
boxplot(tempurature~season,ylab="Tempurature",xlab= "Seasons")
```

Console

 boxplot(tempurature, formula=season)
> boxplot(tempurature, formula=season="winter")
Error: unexpected '=' in "boxplot(tempurature, formula=season="
> boxplot(tempurature, formula=season=="winter")
> boxplot(tempurature, formula=(season=="winter"))
> boxplot(tempurature, formula=tempurature~season)
> ?formula
> ?boxplot
> boxplot(x=tempurature, formula=tempurature~season)
> ?~
+ 

> View(spore_disp)
> boxplot(tempurature~season)
> boxplot(tempurature~season,ylab="Tempurature",xlab= "Seasons")


**Question 8:**

If it's alright I think I'm going to try an format my research code into RMarkdown and edit it into different graph styles in a seprate Rmarkdown document that I will also send you. Mostly, I just have this huge messy chunk of code that might clutter this lab. 